---
title: "TAX_Compare"
author: "Keyu Nie"
date: "11/4/2017"
output: html_document
runtime: shiny
---
#1 New Tax Rate
-------
The House plan creates the following tax chart.

**Income Tax Rate** ||**Income Levels for Those Filing As:**| 
:-----:|:-----:|:-----:|:-----:
Current|Tax Act|Single|Married-Joint
10 - 15%|12%|$0-$44,999|$0 - $89,999
25 - 28%|25%|$45,000 - $199,999 |$90,000 - $259,999
28 - 39.6%|35%|$200,000 - $499,999|$260,000 - $999,999
39.60%|39.60%|$500,000+|$1M+

```{r, include=FALSE}
library(ggplot2)
#Here taxble_income a tax term.Return the old way of tax calculation (not consider AMT) 
tax_calculator<-function(taxable_income, bracket, rate){
  sum(diff(pmin(taxable_income, bracket)) * rate)
}
old_tax<-function(taxable_income, joint = TRUE){ 
rate= c(0.1, 0.15, 0.25, 0.28, 0.33, 0.35, 0.396)
if (joint){
  bracket<-c(0, 18650, 75900, 153100, 233350, 416700,  470700, Inf)
}
else{
  bracket<-c(0, 9325, 37950, 91900, 191650,  416700,  418400, Inf)
}
return(tax_calculator(taxable_income, bracket, rate))
}
```
```{r, include=FALSE}
#Return the new way of tax calculation
new_tax<-function(taxable_income, joint = TRUE){ 
rate= c(0.12, 0.25, 0.35, 0.396)
if (joint){
  bracket<-c(0, 90000, 260000, 1000000,Inf)
}else{
  bracket<-c(0, 45000, 200000, 500000,Inf)
}
return(tax_calculator(taxable_income, bracket, rate))
}
```
#2 Remove AMT
The Alternative Minimum Tax exemption amount for tax year 2017 is $54,300 ($84,500 for married couples filing jointly) and begins to phase out at $120,700 ($160,900 for married couples filing jointly). For tax year 2017, the 26 percent and 28 percent tax rate applies to taxpayers with taxable incomes lower and above $187,800.
```{r, include=FALSE}
AMT_tax<- function(taxable_income, joint=TRUE){
if (joint){
    amt=max(0, 84500 - max(taxable_income - 160900, 0)/4)
  } else {
    amt=max(0, 54300 - max(taxable_income - 120700, 0)/4)
  }
rate=c(0.26, 0.28)
bracket=c(0, 187800,Inf)
return(max(sum(diff(pmin(taxable_income-amt,bracket))*rate),0))
}
```
#3 Remove Personal Exemption. 
#4 Double Standard Deduction.
#5 Cap Tax Deduction upto $10,000. Mortgage Interests Deduction upto $500,000 instead of $1,000,000.

```{r, include=FALSE}

MTG_IR_deduction_old<-function(loan_amount, interest_rate){
  return(min(1000000, loan_amount)*interest_rate)
}
MTG_IR_deduction_new<-function(loan_amount, interest_rate,affected_status=TRUE){
  if(affected_status)
  {return(min(500000, loan_amount)*interest_rate)}
  else{return(min(1000000, loan_amount)*interest_rate)}
}

tax_paying_deduction_old<-function(gross_income, efficient_state_rate, local_tax)
  {return(gross_income*efficient_state_rate + local_tax)}

tax_paying_deduction_new<-function(gross_income, efficient_state_rate, local_tax)
{return(min(10000.0, gross_income*efficient_state_rate + local_tax))}

PersonalExemption_deduction_old<-function(gross_income, child, joint = TRUE)
{  if (joint)
  {member=2+child
  phaseout = min(0.02*round((max(gross_income - 311300, 0)/2500 + 1e-7)), 1)
 return(4050*member*(1 - phaseout))}
 else
  {member=1+child
    phaseout = min(0.02*round(max(gross_income - 259000, 0)/2500 + 1e-7), 1)
return(4050*member*(1 - phaseout))}
}
PersonalExemption_deduction_new<-0

ChildCare_Credit_old<-function(gross_income, child, joint = TRUE)
  {if(joint)
  {phaseout = round(max(gross_income - 110000, 0)/20)
return(max(0,1000*child -  phaseout))
}
else
  {phaseout = round(max(gross_income - 55000, 0)/20)
return(max(0,1000*child -  phaseout))}
}

ChildCare_Credit_new<-function(gross_income, child, joint = TRUE)
{ if(joint)
  {
  phaseout = round(max(gross_income - 230000, 0)/20)
return(max(0,1600*child -  phaseout))
}
else{
  phaseout = round(max(gross_income - 115000, 0)/20)
return(max(0,1600*child -  phaseout))}
}
Standard_deduct_old<-function(joint = TRUE){ if(joint)
  {
return(12600)
}
else{
return(6300)}
}
Standard_deduct_new<-function(joint = TRUE){ if(joint)
  {
return(24000)
}
else{
return(12000)}
}

Itemized<-function(AGI,itemized_amt,joint = TRUE)
{
  if(joint){
    item_deduct=itemized_amt-max(min((AGI-313800)*0.03, 0.8*itemized_amt),0)
    return(item_deduct)
  }
  else{
    item_deduct=itemized_amt-max(min((AGI-261500)*0.03, 0.8*itemized_amt),0)
    return(item_deduct)
  }
}
```

```{r, echo=FALSE}

inputPanel(
  selectInput("joint1", label = "Married couples filing jointly?",
              choices = c(TRUE, FALSE), selected = TRUE),
  numericInput("state_rate", "State tax rate", 0.08, min = 0, max = 1),
  numericInput("Loan_Amount", "Mortgage Amount", 1000000, min = 0, max = 10000000),
  numericInput("interest_rate", "Loan Interest tax rate", 0.04, min = 0, max = 1),
  numericInput("local_tax", "Property Tax Amount", 18000, min = 0, max = 10000000),
  selectInput("child", label = "Number of Child",
              choices = c(0,1,2,3,4,5), selected = 2),
  sliderInput("income", label = "W2 Gross Income Range",
              min = 0, max = 1000000, value = 500000, step = 100000),
  numericInput("pre_tax_deduct", "  Qualified plan contributions", 18000, min = 0, max = 10000000),
  numericInput("Salary", "Your Salary", 150000, min = 0, max = 10000000),
  selectInput("affected_status", label = "Loan after Nov/02/2017?",choices = c(TRUE, FALSE), selected = TRUE)
)
```
**Note here "Married couples filing jointly?" FALSE means filling as single**

```{r, echo=FALSE}
renderPlot({
  gross_income=seq(0,input$income,by=500)
len=length(gross_income)
the_old_tax=the_new_tax=c()
#joint1=TRUE
child=as.numeric(input$child)
#interest_rate=0.04
#Loan_Amount=0
#state_rate=0.08
#pre_tax_deduct=0
#local_tax=0
for (i in 1:len){
  standard_deduct_old=Standard_deduct_old(joint = input$joint1)
  standard_deduct_new=Standard_deduct_new(joint = input$joint1)
  childcare_credit_old=ChildCare_Credit_old(gross_income[i], child, joint = input$joint1)
  childcare_credit_new=ChildCare_Credit_new(gross_income[i], child, joint = input$joint1)
  AGI=gross_income[i]-input$pre_tax_deduct
  Personal_exemptions_old=PersonalExemption_deduction_old(gross_income[i], child, joint = input$joint1)
  Personal_exemptions_new=PersonalExemption_deduction_new
  mortgage_deduction_old=MTG_IR_deduction_old(input$Loan_Amount, input$interest_rate)
  mortgage_deduction_new=MTG_IR_deduction_new(input$Loan_Amount, input$interest_rate, input$affected_status)
  tax_deduction_old=tax_paying_deduction_old(gross_income[i], input$state_rate, input$local_tax)
  tax_deduction_new=tax_paying_deduction_new(gross_income[i], input$state_rate, input$local_tax)
  itemized_deduction_old=tax_deduction_old+mortgage_deduction_old+Personal_exemptions_old
  itemized_deduction_new=tax_deduction_new+mortgage_deduction_new+Personal_exemptions_new
  itemized_deduction_old_revised=Itemized(AGI=AGI,itemized_deduction_old, input$joint1)
  itemized_deduction_new_revised=Itemized(AGI=AGI,itemized_deduction_new, input$joint1)
  taxable_income_old=AGI-max(standard_deduct_old, itemized_deduction_old_revised)
  taxable_income_new=AGI-max(standard_deduct_new, itemized_deduction_new_revised)
  taxable_income_amt=AGI-mortgage_deduction_old
  tax_liability_before_credit_old_raw=old_tax(taxable_income_old,joint = input$joint1)
  tax_liability_before_credit_amt=AMT_tax(taxable_income_amt,joint = input$joint1)
  tax_liability_before_credit_old=max(tax_liability_before_credit_old_raw, tax_liability_before_credit_amt)
  tax_liability_before_credit_new=new_tax(taxable_income_new,joint = input$joint1)
  tax_liability_old=tax_liability_before_credit_old-childcare_credit_old
  tax_liability_new=tax_liability_before_credit_new-childcare_credit_new
  the_old_tax=c(the_old_tax,tax_liability_old)
  the_new_tax=c(the_new_tax,tax_liability_new)
}
df_2 <- data.frame(GROSS=rep(gross_income, 2), tax=c(the_old_tax,the_new_tax), 
                 variable=rep(c("old_tax", "new_tax"), each=len))
  ggplot(data = df_2, aes(x=GROSS, y=tax)) + geom_line(aes(colour=variable))+ ggtitle("Plot new and old tax calulation.") +
  xlab("Gross Income") + ylab("Tax Liable Amount")
})

```

```{r, echo=FALSE}
  renderPlot({
      gross_income=seq(0,input$income,by=500)
len=length(gross_income)
the_old_tax=the_new_tax=c()
#joint1=TRUE
child=as.numeric(input$child)
#interest_rate=0.04
#Loan_Amount=0
#state_rate=0.08
#pre_tax_deduct=0
#local_tax=0
for (i in 1:len){
  standard_deduct_old=Standard_deduct_old(joint = input$joint1)
  standard_deduct_new=Standard_deduct_new(joint = input$joint1)
  childcare_credit_old=ChildCare_Credit_old(gross_income[i], child, joint = input$joint1)
  childcare_credit_new=ChildCare_Credit_new(gross_income[i], child, joint = input$joint1)
  AGI=gross_income[i]-input$pre_tax_deduct
  Personal_exemptions_old=PersonalExemption_deduction_old(gross_income[i], child, joint = input$joint1)
  Personal_exemptions_new=PersonalExemption_deduction_new
  mortgage_deduction_old=MTG_IR_deduction_old(input$Loan_Amount, input$interest_rate)
  mortgage_deduction_new=MTG_IR_deduction_new(input$Loan_Amount, input$interest_rate, input$affected_status)
  tax_deduction_old=tax_paying_deduction_old(gross_income[i], input$state_rate, input$local_tax)
  tax_deduction_new=tax_paying_deduction_new(gross_income[i], input$state_rate, input$local_tax)
  itemized_deduction_old=tax_deduction_old+mortgage_deduction_old+Personal_exemptions_old
  itemized_deduction_new=tax_deduction_new+mortgage_deduction_new+Personal_exemptions_new
  itemized_deduction_old_revised=Itemized(AGI=AGI,itemized_deduction_old, input$joint1)
  itemized_deduction_new_revised=Itemized(AGI=AGI,itemized_deduction_new, input$joint1)
  taxable_income_old=AGI-max(standard_deduct_old, itemized_deduction_old_revised)
  taxable_income_new=AGI-max(standard_deduct_new, itemized_deduction_new_revised)
  taxable_income_amt=AGI-mortgage_deduction_old
  tax_liability_before_credit_old_raw=old_tax(taxable_income_old,joint = input$joint1)
  tax_liability_before_credit_amt=AMT_tax(taxable_income_amt,joint = input$joint1)
  tax_liability_before_credit_old=max(tax_liability_before_credit_old_raw, tax_liability_before_credit_amt)
  tax_liability_before_credit_new=new_tax(taxable_income_new,joint = input$joint1)
  tax_liability_old=tax_liability_before_credit_old-childcare_credit_old
  tax_liability_new=tax_liability_before_credit_new-childcare_credit_new
  the_old_tax=c(the_old_tax,tax_liability_old)
  the_new_tax=c(the_new_tax,tax_liability_new)
}
    New_minus_Old=the_new_tax-the_old_tax
    plot(x=gross_income, y=New_minus_Old, xlab="Gross Income", ylab="Tax difference New-Old", main='Plot the difference between new tax proposal and old tax')

})
```

```{r, echo=FALSE}
 renderText({
gross_income=input$Salary
#the_old_tax=the_new_tax=c()
#joint1=TRUE
  child=as.numeric(input$child)
  standard_deduct_old=Standard_deduct_old(joint = input$joint1)
  standard_deduct_new=Standard_deduct_new(joint = input$joint1)
  childcare_credit_old=ChildCare_Credit_old(gross_income, child, joint = input$joint1)
  childcare_credit_new=ChildCare_Credit_new(gross_income, child, joint = input$joint1)
  AGI=gross_income-input$pre_tax_deduct
  Personal_exemptions_old=PersonalExemption_deduction_old(gross_income, child, joint = input$joint1)
  Personal_exemptions_new=PersonalExemption_deduction_new
  mortgage_deduction_old=MTG_IR_deduction_old(input$Loan_Amount, input$interest_rate)
  mortgage_deduction_new=MTG_IR_deduction_new(input$Loan_Amount, input$interest_rate, input$affected_status)
  tax_deduction_old=tax_paying_deduction_old(gross_income, input$state_rate, input$local_tax)
  tax_deduction_new=tax_paying_deduction_new(gross_income, input$state_rate, input$local_tax)
  itemized_deduction_old=tax_deduction_old+mortgage_deduction_old+Personal_exemptions_old
  itemized_deduction_new=tax_deduction_new+mortgage_deduction_new+Personal_exemptions_new
  itemized_deduction_old_revised=Itemized(AGI=AGI,itemized_deduction_old, input$joint1)
  itemized_deduction_new_revised=Itemized(AGI=AGI,itemized_deduction_new, input$joint1)
  taxable_income_old=AGI-max(standard_deduct_old, itemized_deduction_old_revised)
  taxable_income_new=AGI-max(standard_deduct_new, itemized_deduction_new_revised)
  taxable_income_amt=AGI-mortgage_deduction_old
  tax_liability_before_credit_old_raw=old_tax(taxable_income_old,joint = input$joint1)
  tax_liability_before_credit_amt=AMT_tax(taxable_income_amt,joint = input$joint1)
  tax_liability_before_credit_old=max(tax_liability_before_credit_old_raw, tax_liability_before_credit_amt)
  tax_liability_before_credit_new=new_tax(taxable_income_new,joint = input$joint1)
  tax_liability_old=tax_liability_before_credit_old-childcare_credit_old
  tax_liability_new=tax_liability_before_credit_new-childcare_credit_new
#  the_old_tax=c(the_old_tax,tax_liability_old)
#  the_new_tax=c(the_new_tax,tax_liability_new)
#    New_minus_Old=the_new_tax-the_old_tax
str1=paste0("Old Standard Tax is ",tax_liability_before_credit_old_raw, ". ")
str2=paste0("Old AMT Tax is ",tax_liability_before_credit_amt, ". ")
str3=paste0("New Standard Tax is ",tax_liability_before_credit_new, ". ")
str4=paste0("Old Taxable Income is ",taxable_income_old, ". ")
str5=paste0("New Taxable Income is ",taxable_income_new, ". ")
str6=paste0("Old Itemized Deduction is ",itemized_deduction_old_revised, ". ")
str7=paste0("New Itemized Deduction is ",itemized_deduction_new_revised, ". ")

HTML(paste(str1, str2, str3, str4,str5, str6,str7, sep='\n'))
})
```



